### 2021-11-30 Artificial Intelligence and Machine Learning/PADL_conference/bbl.tex
pulled references inline

\begin{thebibliography}{10}
\providecommand{\url}[1]{\texttt{#1}}
\providecommand{\urlprefix}{URL }
\providecommand{\doi}[1]{https://doi.org/#1}

\bibitem{abgemuscwo15c}
Abseher, M., Gebser, M., Musliu, N., Schaub, T., Woltran, S.: Shift design with
  answer set programming. Fundamenta Informaticae  \textbf{147}(1),  1--25
  (2016)

\bibitem{adibi2014clustering}
Adibi, M., Shahrabi, J.: A clustering-based modified variable neighborhood
  search algorithm for a dynamic job shop scheduling problem. International
  Journal of Advanced Manufacturing Technology  \textbf{70}(9-12),  1955--1961
  (2014)

\bibitem{al2017job}
Al-Ashhab, M., Munshi, S., Oreijah, M., Ghulman, H.: Job shop scheduling using
  mixed integer programming. International Journal Of Modern Engineering
  Research  \textbf{7}(3),  2:23--2:29 (2017)

\bibitem{baker1974introduction}
Baker, K.: Introduction to Sequencing and Scheduling. John Wiley \& Sons (1974)

\bibitem{balduccini11a}
Balduccini, M.: Industrial-size scheduling with {ASP}+{CP}. In: Proceedings of
  the Eleventh International Conference on Logic Programming and Nonmonotonic
  Reasoning (LPNMR'11), pp. 284--296. Springer-Verlag (2011)

\bibitem{bainkaokscsotawa18a}
Banbara, M., Inoue, K., Kaufmann, B., Okimoto, T., Schaub, T., Soh, T., Tamura,
  N., Wanko, P.: teaspoon: Solving the curriculum-based course timetabling
  problems with answer set programming. Annals of Operations Research
  \textbf{275}(1),  3--37 (2019)

\bibitem{bengio2020machine}
Bengio, Y., Lodi, A., Prouvost, A.: Machine learning for combinatorial
  optimization: A methodological tour d'horizon. European Journal of
  Operational Research  \textbf{290}(2),  405--421 (2021)

\bibitem{blackstone1982state}
Blackstone, J., Phillips, D., Hogg, G.: A state-of-the-art survey of
  dispatching rules for manufacturing job shop operations. International
  Journal of Production Research  \textbf{20}(1),  27--45 (1982)

\bibitem{DBLP:journals/jim/CalisB15}
{\c{C}}alis, B., Bulkan, S.: A research survey: Review of {AI} solution
  strategies of job shop scheduling problem. Journal of Intelligent
  Manufacturing  \textbf{26}(5),  961--973 (2015)

\bibitem{chaudhry2016research}
Chaudhry, I., Khan, A.: A research survey: Review of flexible job shop
  scheduling techniques. International Transactions in Operational Research
  \textbf{23}(3),  551--591 (2016)

\bibitem{coltep19a}
{Da Col}, G., Teppan, E.: Industrial size job shop scheduling tackled by
  present day {CP} solvers. In: Proceedings of the Twenty-fifth International
  Conference on Principles and Practice of Constraint Programming (CP'19), pp.
  144--160. Springer-Verlag (2019)

\bibitem{demirbilek2019dynamically}
Demirbilek, M., Branke, J., Strauss, A.: Dynamically accepting and scheduling
  patients for home healthcare. Health Care Management Science  \textbf{22}(1),
   140--155 (2019)

\bibitem{demirkol1998benchmarks}
Demirkol, E., Mehta, S., Uzsoy, R.: Benchmarks for shop scheduling problems.
  European Journal of Operational Research  \textbf{109}(1),  137--141 (1998)

\bibitem{ding2020unified}
Ding, H., Xu, J.: A unified framework for clustering constrained data without
  locality property. Algorithmica  \textbf{82}(4),  808--852 (2020)

\bibitem{dogagrmamopo21a}
Dodaro, C., Galat{\`{a}}, G., Grioni, A., Maratea, M., Mochi, M., Porro, I.: An
  {ASP}-based solution to the chemotherapy treatment scheduling problem. Theory
  and Practice of Logic Programming  \textbf{First
  View}(\url{https://doi.org/10.1017/S1471068421000363}),  1--17 (2021)

\bibitem{el2020job}
El-Kholany, M., Gebser, M.: Job shop scheduling with multi-shot {ASP}. In:
  Proceedings of the Fourth Workshop on Trends and Applications of Answer Set
  Programming (TAASP'20).
  \url{http://www.kr.tuwien.ac.at/events/taasp20/papers/TAASP_2020_paper_4.pdf}
  (2020)

\bibitem{Forgy1965ClusterAO}
Forgy, E.: Cluster analysis of multivariate data: Efficiency versus
  interpretability of classifications. Biometrics  \textbf{21},  768--769
  (1965)

\bibitem{frscel21a}
Francescutto, G., Schekotihin, K., El{-}Kholany, M.: Solving a multi-resource
  partial-ordering flexible variant of the job-shop scheduling problem with
  hybrid {ASP}. In: Proceedings of the Seventeenth European Conference on
  Logics in Artificial Intelligence (JELIA'21), pp. 313--328. Springer-Verlag
  (2021)

\bibitem{french82a}
French, S.: Sequencing and Scheduling: An Introduction to the Mathematics of
  the Job-shop. John Wiley \& Sons (1982)

\bibitem{garey1976complexity}
Garey, M., Johnson, D., Sethi, R.: The complexity of flowshop and jobshop
  scheduling. Mathematics of Operations Research  \textbf{1}(2),  117--129
  (1976)

\bibitem{gekakasc17a}
Gebser, M., Kaminski, R., Kaufmann, B., Schaub, T.: Multi-shot {ASP} solving
  with clingo. Theory and Practice of Logic Programming  \textbf{19}(1),
  27--82 (2019)

\bibitem{harrath2002genetic}
Harrath, Y., Chebel-Morello, B., Zerhouni, N.: A genetic algorithm and data
  mining based meta-heuristic for job shop scheduling problem. In: Proceedings
  of the IEEE International Conference on Systems, Man and Cybernetics
  (SMC'02). IEEE (2002)

\bibitem{ismail2012production}
Ismail, R., Othman, Z., Bakar, A.: A production schedule generator framework
  for pattern sequential mining. In: Proceedings of the Seventh International
  Conference on Computing and Convergence Technology (ICCCT'12), pp. 784--788.
  IEEE (2012)

\bibitem{janakbhai2021blockchain}
Janakbhai, N., Saurin, M., Patel, M.: Blockchain-based intelligent
  transportation system with priority scheduling. In: Data Science and
  Intelligent Applications, pp. 311--317. Springer-Verlag (2021)

\bibitem{jones1998survey}
Jones, A., Rabelo, L., Sharawi, A.: Survey of job shop scheduling techniques.
  National Institute of Standards and Technology  \textbf{Encyclopedia of
  Electrical and Electronics
  Engineering}(\url{https://tsapps.nist.gov/publication/get_pdf.cfm?pub_id=821200})
  (1998)

\bibitem{koonce2000using}
Koonce, D., Tsai, S.: Using data mining to find patterns in genetic algorithm
  solutions to a job shop schedule. Computers \& Industrial Engineering
  \textbf{38}(3),  361--374 (2000)

\bibitem{lenstra1979computational}
Lenstra, J., Rinnooy~Kan, A.: Computational complexity of discrete optimization
  problems. Annals of Discrete Mathematics  \textbf{4},  121--140 (1979)

\bibitem{meng2020mixed}
Meng, L., Zhang, C., Ren, Y., Zhang, B., Lv, C.: Mixed-integer linear
  programming and constraint programming formulations for solving distributed
  flexible job shop scheduling problem. Computers \& Industrial Engineering
  \textbf{142},  Article 106347 (2020)

\bibitem{nasiri2019data}
Nasiri, M., Salesi, S., Rahbari, A., Meydani, N., Abdollai, M.: A data mining
  approach for population-based methods to solve the {JSSP}. Soft Computing
  \textbf{23}(21),  11107--11122 (2019)

\bibitem{nouiri2018effective}
Nouiri, M., Bekrar, A., Jemai, A., Niar, S., Ammari, A.: An effective and
  distributed particle swarm optimization algorithm for flexible job-shop
  scheduling problem. Journal of Intelligent Manufacturing  \textbf{29}(3),
  603--615 (2018)

\bibitem{ovacik2012decomposition}
Ovacik, I., Uzsoy, R.: Decomposition Methods for Complex Factory Scheduling
  Problems. Kluwer Academic Publishers (1997)

\bibitem{pezzella2008genetic}
Pezzella, F., Morganti, G., Ciaschetti, G.: A genetic algorithm for the
  flexible job-shop scheduling problem. Computers \& Operations Research
  \textbf{35}(10),  3202--3212 (2008)

\bibitem{rigralmaliiile12a}
Ricca, F., Grasso, G., Alviano, M., Manna, M., Lio, V., Iiritano, S., Leone,
  N.: Team-building with answer set programming in the {G}ioia-{T}auro seaport.
  Theory and Practice of Logic Programming  \textbf{12}(3),  361--381 (2012)

\bibitem{schoenfelder2020nurse}
Schoenfelder, J., Bretthauer, K., Wright, D., Coe, E.: Nurse scheduling with
  quick-response methods: Improving hospital performance, nurse workload, and
  patient experience. European Journal of Operational Research
  \textbf{283}(1),  390--403 (2020)

\bibitem{shahzad2010discovering}
Shahzad, A., Mebarki, N.: Discovering dispatching rules for job shop scheduling
  problem through data mining. In: Proceedings of the Eighth International
  Conference of Modeling and Simulation (MOSIM'10).
  \url{https://www.academia.edu/3068769/DISCOVERING_DISPATCHING_RULES_FOR_JOB_SHOP_SCHEDULING_PROBLEM_THROUGH_DATA_MINING}
  (2010)

\bibitem{singer2001decomposition}
Singer, M.: Decomposition methods for large job shops. Computers \& Operations
  Research  \textbf{28}(3),  193--207 (2001)

\bibitem{SOTSKOV1995237}
Sotskov, Y., Shakhlevich, N.: {NP}-hardness of shop-scheduling problems with
  three jobs. Discrete Applied Mathematics  \textbf{59}(3),  237--266 (1995)

\bibitem{taillard1993benchmarks}
Taillard, E.: Benchmarks for basic scheduling problems. European Journal of
  Operational Research  \textbf{64}(2),  278--285 (1993)

\bibitem{tagesc21a}
Tassel, P., Gebser, M., Schekotihin, K.: A reinforcement learning environment
  for job-shop scheduling. In: Proceedings of the ICAPS 2021 Workshop on
  Planning and Reinforcement Learning (PRL'21).
  \url{https://prl-theworkshop.github.io/prl2021/papers/PRL2021_paper_9.pdf}
  (2021)

\bibitem{uzsoy2000performance}
Uzsoy, R., Wang, C.: Performance of decomposition procedures for job shop
  scheduling problems with bottleneck machines. International Journal of
  Production Research  \textbf{38}(6),  1271--1286 (2000)

\bibitem{DBLP:journals/informs/VaessensAL96}
Vaessens, R., Aarts, E., Lenstra, J.: Job shop scheduling by local search.
  INFORMS Journal on Computing  \textbf{8}(3),  302--317 (1996)

\bibitem{wagstaff2001constrained}
Wagstaff, K., Cardie, C., Rogers, S., Schr{\"{o}}dl, S.: Constrained k-means
  clustering with background knowledge. In: Proceedings of the Eighteenth
  International Conference on Machine Learning (ICML'01), pp. 577--584. Morgan
  Kaufmann (2001)

\bibitem{wang2019routing}
Wang, H.: Routing and scheduling for a last-mile transportation system.
  Transportation Science  \textbf{53}(1),  131--147 (2019)

\bibitem{zhai2014decomposition}
Zhai, Y., Liu, C., Chu, W., Guo, R., Liu, C.: A decomposition heuristics based
  on multi-bottleneck machines for large-scale job shop scheduling problems.
  Journal of Industrial Engineering and Management  \textbf{7}(5),  1397--1414
  (2014)

\bibitem{zhang2019framework}
Zhang, H., Basu, S., Davidson, I.: A framework for deep constrained clustering:
  Algorithms and advances. In: Proceedings of the Joint European Conference on
  Machine Learning and Knowledge Discovery in Databases (ECML/PKDD'19), pp.
  57--72. Springer-Verlag (2019)

\bibitem{zhang2010hybrid}
Zhang, R., Wu, C.: A hybrid approach to large-scale job shop scheduling.
  Applied Intelligence  \textbf{32}(1),  47--59 (2010)

\end{thebibliography}